Analysis

### 2022-07-05 03_EDA_CTCFBSDB.Rmd
Add liftOver

# Input files
# Predicted data
fileNameIn1 <- file.path(dir_project, "data/allcomp.txt.gz")
if (!file.exists(fileNameIn1)) {
  download.file("https://insulatordb.uthsc.edu/download/allcomp.txt.gz", fileNameIn1)
}
# Experimental data
fileNameIn2 <- file.path(dir_project, "data/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz")
# download.file("https://insulatordb.uthsc.edu/download/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz", fileNameIn2)
# Output files
fileNameOut1 <- file.path(dir_project, "RData", "hg18.CTCFBSDB.CTCF_predicted_human.RData")
fileNameOut1.1 <- file.path(dir_project, "RData", "hg18.CTCFBSDB.CTCF_predicted_human.bed")
fileNameOut2 <- file.path(dir_project, "RData", "mm8.CTCFBSDB.CTCF_predicted_mouse.RData")
fileNameOut2.1 <- file.path(dir_project, "RData", "mm8.CTCFBSDB.CTCF_predicted_mouse.bed")
# fileNameOut3 <- file.path(dir_project, "results/hg19.CTCFBSDB.CTCF_experimental_human.RData")
# fileNameOut3.1 <- file.path(dir_project, "results/hg19.CTCFBSDB.CTCF_experimental_human.bed")
# fileNameOut4 <- file.path(dir_project, "results/mm9.CTCFBSDB.CTCF_experimental_mouse.RData")
# fileNameOut4.1 <- file.path(dir_project, "results/mm9.CTCFBSDB.CTCF_experimental_mouse.bed")
```

@@

>>>>

# Input files
# Predicted data
fileNameIn1 <- file.path(dir_project, "data/allcomp.txt.gz")
if (!file.exists(fileNameIn1)) {
  download.file("https://insulatordb.uthsc.edu/download/allcomp.txt.gz", fileNameIn1)
}
# Experimental data
fileNameIn2 <- file.path(dir_project, "data/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz")
# download.file("https://insulatordb.uthsc.edu/download/CTCFBSDB_all_exp_sites_Sept12_2012.txt.gz", fileNameIn2)
# Output files
fileNameOut1 <- file.path(dir_project, "RData", "hg18.CTCFBSDB.CTCF_predicted_human.RData")
fileNameOut1.1 <- file.path(dir_project, "RData", "hg18.CTCFBSDB.CTCF_predicted_human.bed")
fileNameOut2 <- file.path(dir_project, "RData", "mm8.CTCFBSDB.CTCF_predicted_mouse.RData")
fileNameOut2.1 <- file.path(dir_project, "RData", "mm8.CTCFBSDB.CTCF_predicted_mouse.bed")
# liftOver human
fileNameOut3 <- file.path(dir_project, "RData", "hg19.CTCFBSDB.CTCF_predicted_human.RData")
fileNameOut3.1 <- file.path(dir_project, "RData", "hg19.CTCFBSDB.CTCF_predicted_human.bed")
fileNameOut4 <- file.path(dir_project, "RData", "hg38.CTCFBSDB.CTCF_predicted_human.RData")
fileNameOut4.1 <- file.path(dir_project, "RData", "hg38.CTCFBSDB.CTCF_predicted_human.bed")
# liftOver mouse
fileNameOut5 <- file.path(dir_project, "RData", "mm9.CTCFBSDB.CTCF_predicted_mouse.RData")
fileNameOut5.1 <- file.path(dir_project, "RData", "mm9.CTCFBSDB.CTCF_predicted_mouse.bed")
fileNameOut6 <- file.path(dir_project, "RData", "mm10.CTCFBSDB.CTCF_predicted_mouse.RData")
fileNameOut6.1 <- file.path(dir_project, "RData", "mm10.CTCFBSDB.CTCF_predicted_mouse.bed")
```

@@

# LiftOver

```{r}
# Function to download a chain for the gf18_to_XX conversion and create a GRanges object
liftOver_custom <- function(URL = "https://hgdownload.cse.ucsc.edu/goldenpath/hg18/liftOver/hg18ToHg19.over.chain.gz", gr_in = hg18.CTCFBSDB.CTCF_predicted_human, genome_id = "hg19"){
  fileChain <- file.path(dir_project, "data", basename(URL))
  if (!file.exists(str_remove(fileChain, ".gz"))) {
    download.file(URL, destfile = fileChain)
    R.utils::gunzip(fileChain)
  }

  ch <- import.chain(str_remove(fileChain, ".gz"))
  gr_converted <- liftOver(gr_in, ch) %>% unlist()
  # Get chromosome info and match it to the chromosome order in ctcfBED
  chrom_data <- GenomeInfoDb::getChromInfoFromUCSC(genome = genome_id)
  # Subset to common chromosomes
  common_chromosomes <- intersect(chrom_data$chrom, seqlevels(gr_converted))
  chrom_data <- chrom_data[chrom_data$chrom %in% common_chromosomes, ]
  gr_converted <- keepSeqlevels(gr_converted, common_chromosomes, pruning.mode = "tidy")
  # Match order
  chrom_data <- chrom_data[match(seqlevels(gr_converted), chrom_data$chrom), ]
  # Check if chromosome order is the same
  if (!all.equal(seqlevels(gr_converted), chrom_data$chrom)) {
    print(paste("Chromosome order does not match for", genome_id, "genome."))
    break
  }
  # Assign seqinfo data
  seqlengths(gr_converted) <- chrom_data$size
  isCircular(gr_converted) <- chrom_data$circular
  genome(gr_converted)     <- genome_id
  return(gr_converted)  
}

# hg18.CTCFBSDB.CTCF_predicted_human to hg19 conversion
gr_out <- "hg19.CTCFBSDB.CTCF_predicted_human"
# Assign the converted GRanges to the variable name
assign(gr_out, liftOver_custom(URL = "https://hgdownload.cse.ucsc.edu/goldenpath/hg18/liftOver/hg18ToHg19.over.chain.gz", gr_in = hg18.CTCFBSDB.CTCF_predicted_human, genome_id = "hg19"))
# Save as RData object. subdir is the character name of the CTCF GRanges variable
save(list = gr_out, file = fileNameOut3)
# load(fileNameOut4)
# Save BED files
export.bed(hg19.CTCFBSDB.CTCF_predicted_human, fileNameOut3.1)

# hg18.CTCFBSDB.CTCF_predicted_human to hg38 conversion
gr_out <- "hg38.CTCFBSDB.CTCF_predicted_human"
# Assign the converted GRanges to the variable name
assign(gr_out, liftOver_custom(URL = "https://hgdownload.cse.ucsc.edu/goldenpath/hg18/liftOver/hg18ToHg38.over.chain.gz", gr_in = hg18.CTCFBSDB.CTCF_predicted_human, genome_id = "hg38"))
# Save as RData object. subdir is the character name of the CTCF GRanges variable
save(list = gr_out, file = fileNameOut4)
# load(fileNameOut4)
# Save BED files
export.bed(hg38.CTCFBSDB.CTCF_predicted_human, fileNameOut4.1)

# mm8.CTCFBSDB.CTCF_predicted_mouse to mm9 conversion
gr_out <- "mm9.CTCFBSDB.CTCF_predicted_mouse"
# Assign the converted GRanges to the variable name
assign(gr_out, liftOver_custom(URL = "https://hgdownload.cse.ucsc.edu/goldenpath/mm8/liftOver/mm8ToMm9.over.chain.gz", gr_in = mm8.CTCFBSDB.CTCF_predicted_mouse, genome_id = "mm9"))
# Save as RData object. subdir is the character name of the CTCF GRanges variable
save(list = gr_out, file = fileNameOut5)
# load(fileNameOut4)
# Save BED files
export.bed(mm9.CTCFBSDB.CTCF_predicted_mouse, fileNameOut5.1)

# mm8.CTCFBSDB.CTCF_predicted_mouse to mm10 conversion
gr_out <- "mm10.CTCFBSDB.CTCF_predicted_mouse"
# Assign the converted GRanges to the variable name
assign(gr_out, liftOver_custom(URL = "https://hgdownload.cse.ucsc.edu/goldenpath/mm8/liftOver/mm8ToMm10.over.chain.gz", gr_in = mm8.CTCFBSDB.CTCF_predicted_mouse, genome_id = "mm10"))
# Save as RData object. subdir is the character name of the CTCF GRanges variable
save(list = gr_out, file = fileNameOut6)
# load(fileNameOut4)
# Save BED files
export.bed(mm9.CTCFBSDB.CTCF_predicted_mouse, fileNameOut6.1)
```
